\singlespacing \vspace{-1.5cm} \LARGE Trade-offs among plant reproductive traits determine interactions with floral visitors

### 2021-11-26 Manuscript/Manuscript.Rmd
wording edit

\vspace{-2.2cm}
\singlespacing
\large 
\textbf{Jose B. Lanuza$^{1,2}*$, Romina Rader$^{1}$, Jamie Stavert$^{3}$, Liam K. Kendall$^{4}$, Manu E. Saunders$^{1}$ and Ignasi Bartomeus$^{2}$}

@@

__Plant life strategies are constrained by cost-benefit trade-offs that determine plant form and function. However, despite recent advances in the understanding of trade-offs for vegetative and physiological traits, little is known about plant reproductive economics and how they constrain plant life strategies and shape interactions with floral visitors. Here, we investigate plant reproductive trade-offs and how these reproductive compromises drive interactions with floral visitors using a dataset of 17 reproductive traits for 1,506 plant species from 28 plant-pollinator studies across 18 countries. We tested whether a plant’s reproductive strategy predicts its interactions with floral visitors and if the different reproductive traits predict the plant’s role within the pollination network. We found that over half of all plant reproductive trait variation was explained by two independent axes that encompassed plant form and function. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size, while the second axis indicated a pollinator dependency trade-off. Plant reproductive trade-offs somewhat explained the presence or absence of interactions with floral visitors, but not differences in visitation rate. However, we did find important differences in the interaction level among floral visitor guilds on the different axes of trait variation. Finally, we found that plant size and floral rewards were the most important traits in the understanding of the species network role. Our results highlight the importance of plant reproductive trade-offs in determining plant life strategies and plant-pollinator interactions in a global context.__

@@

Here, we show that plant species exhibit clear trade-offs among their vegetative and reproductive traits and that these trade-offs determine interactions with floral visitors. These trade-offs are differentiated along three axes of trait variation: (i) flower number - flower size, (ii) pollinator dependence and (iii) style length. These reproductive trade-offs helped explain the presence of floral visitor interactions, but not their visitation rates. Floral visitor guilds formed distinct relationships with the main axes of trait variation. Moreover, we found that the plant species functional roles within pollination networks were best explained by plant size and floral reward related traits.

>>>>

\vspace{-1.8cm}
\singlespacing
\large 
\textbf{Jose B. Lanuza$^{1,2}*$, Romina Rader$^{1}$, Jamie Stavert$^{3}$, Liam K. Kendall$^{4}$, Manu E. Saunders$^{1}$ and Ignasi Bartomeus$^{2}$}

@@

__Plant life strategies are constrained by cost-benefit trade-offs that determine plant form and function. However, despite recent advances in the understanding of trade-offs for vegetative and physiological traits, little is known about plant reproductive economics and how they constrain plant life strategies and shape interactions with floral visitors. Here, we investigate plant reproductive trade-offs and how these reproductive compromises drive interactions with floral visitors using a dataset of 17 reproductive traits for 1,506 plant species from 28 plant-pollinator studies across 18 countries. We tested whether a plant’s reproductive strategy predicts its interactions with floral visitors and if the different reproductive traits predict the plant’s role within the pollination network. We found that over half of all plant reproductive trait variation was explained by two independent axes that encompassed plant form and function. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size, while the second axis indicated a pollinator dependency trade-off. Plant reproductive trade-offs helped explain partly the presence or absence of interactions with floral visitors, but not differences in visitation rate. However, we did find important differences in the interaction level among floral visitor guilds on the different axes of trait variation. Finally, we found that plant size and floral rewards were the most important traits in the understanding of the species network role. Our results highlight the importance of plant reproductive trade-offs in determining plant life strategies and plant-pollinator interactions in a global context.__

@@

Here, we show that plant species exhibit clear trade-offs among their vegetative and reproductive traits and that these trade-offs determine interactions with floral visitors. These trade-offs are differentiated along three axes of trait variation: (i) flower number - flower size, (ii) pollinator dependence and (iii) style length. These reproductive trade-offs helped explain partly the presence of floral visitor interactions, but not their visitation rates. However, floral visitor guilds formed distinct relationships with the main axes of trait variation. Moreover, we found that the plant species functional roles within pollination networks were best explained by plant size and floral reward related traits.

### 2021-11-26 Manuscript/Manuscript.tex
wording edit

\setstretch{2}
\vspace{-2.2cm}
\singlespacing
\large

@@

\textbf{Plant life strategies are constrained by cost-benefit trade-offs that determine plant form and function. However, despite recent advances in the understanding of trade-offs for vegetative and physiological traits, little is known about plant reproductive economics and how they constrain plant life strategies and shape interactions with floral visitors. Here, we investigate plant reproductive trade-offs and how these reproductive compromises drive interactions with floral visitors using a dataset of 17 reproductive traits for 1,506 plant species from 28 plant-pollinator studies across 18 countries. We tested whether a plant's reproductive strategy predicts its interactions with floral visitors and if the different reproductive traits predict the plant's role within the pollination network. We found that over half of all plant reproductive trait variation was explained by two independent axes that encompassed plant form and function. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size, while the second axis indicated a pollinator dependency trade-off. Plant reproductive trade-offs somewhat explained the presence or absence of interactions with floral visitors, but not differences in visitation rate. However, we did find important differences in the interaction level among floral visitor guilds on the different axes of trait variation. Finally, we found that plant size and floral rewards were the most important traits in the understanding of the species network role. Our results highlight the importance of plant reproductive trade-offs in determining plant life strategies and plant-pollinator interactions in a global context.}

@@

Here, we aim to explore the potential trade-offs among reproductive traits and how these influence plant-pollinator interactions. First, we identify the major axes of reproductive trait variation and trade-offs that determine plant form and function. Second, we investigate how plant species' position in trait-space interactions with floral visitors. Finally, we investigate how both the main axes of trait variation, and individual traits, influence plant species' roles within networks using a set of complementary interaction network metrics (i.e., interaction strength, normalized degree and specialization).

>>>>

\setstretch{2}
\vspace{-1.8cm}
\singlespacing
\large

@@

\textbf{Plant life strategies are constrained by cost-benefit trade-offs that determine plant form and function. However, despite recent advances in the understanding of trade-offs for vegetative and physiological traits, little is known about plant reproductive economics and how they constrain plant life strategies and shape interactions with floral visitors. Here, we investigate plant reproductive trade-offs and how these reproductive compromises drive interactions with floral visitors using a dataset of 17 reproductive traits for 1,506 plant species from 28 plant-pollinator studies across 18 countries. We tested whether a plant's reproductive strategy predicts its interactions with floral visitors and if the different reproductive traits predict the plant's role within the pollination network. We found that over half of all plant reproductive trait variation was explained by two independent axes that encompassed plant form and function. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size, while the second axis indicated a pollinator dependency trade-off. Plant reproductive trade-offs helped explain partly the presence or absence of interactions with floral visitors, but not differences in visitation rate. However, we did find important differences in the interaction level among floral visitor guilds on the different axes of trait variation. Finally, we found that plant size and floral rewards were the most important traits in the understanding of the species network role. Our results highlight the importance of plant reproductive trade-offs in determining plant life strategies and plant-pollinator interactions in a global context.}

@@

Here, we aim to explore the potential trade-offs among reproductive traits and how these influence plant-pollinator interactions. First, we identify the major axes of reproductive trait variation and trade-offs that determine plant form and function. Second, we investigate how plant species' position in trait-space influence interactions with floral visitors. Finally, we investigate how both the main axes of trait variation, and individual traits, influence plant species' roles within networks using a set of complementary interaction network metrics (i.e., interaction strength, normalized degree and specialization).
